---
title: "Data Manipulation with 'dplyr'"
output: github_document
date: "2022-09-22"
---

```{r}
knit::opts_chunk$set
collape 
```

```{r}
library (tidyverse)

options(tibble,prin_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)

```

### Select

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

```{r}
select(litters_data, group:gd_of_birth)
```

```{r}
select(litters_data, -pups_survive)

select(litters_data, -pups_survive, -group)
```

# Rename
## new name left, old name right 
```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

## Rename the variable but keep others
```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

```{r}
select(litters_data, starts_with("gd"))
select(litters_data, ends_with("weight"))
```

## Chaning the order

```{r}
select(litters_data, litter_number, pups_survive, everything())
```


# Select rows
`==` distinguish the equal sign of logical
`!-` not equal
`is.na()` to find missing values.
## Test the conndition with `TRUE` adn `FALSE`

* gd_of_birth == 20
* pups_born_alive >= 2
* pups_survive != 4
* !(pups_survive == 4)
* !((pups_survive == 4) & (gd_of_birth == 20))

* group %in% c("Con7", "Con8")
* Select all the row (useful for characters)
* group == "Con7" & gd_of_birth == 20


# Mutation

```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group),
  wt_gain_kg = wt_gain * 2.2
)
```

`wt_gain` create an new dataset 
`str_to_lower(group)` take the string and change to the lowercase


# `arrange`

```{r}
head(arrange(litters_data, group, pups_born_alive), 10)
head(arrange(litters_data, desc(group), pups_born_alive), 10)
```
`head(()),10)` print the fisrt 10 rows 
defult arrangign order small to big


# Piping

## Option 1 - not prefered
```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = 
  mutate(
    litters_data_selected_cols, 
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
litters_data_with_vars_without_missing = 
  drop_na(litters_data_with_vars, wt_gain)
litters_data_with_vars_without_missing
```

## Option 2 - not prefered

```{r}
litters_data_clean = 
  drop_na(
    mutate(
      select(
        janitor::clean_names(
          read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
          ), 
      -pups_survive
      ),
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
    ),
  wt_gain
  )

litters_data_clean
```

## Option 3 - using pipe operator 

### Shortcut Cmd + Shift + M ` %>% `

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)

litters_data
```

read_cvs creates new tibble and pipes into janitor (tibble in, tibble out)
`litters_data=` create one single tibble and save it.
`col_types = "ccddiiii")` identify the variable type.

Need to running chuck everytime when openning the files.


### Fit a linear model (not tidyverse function)
```{r}
litters_data %>%
  lm(wt_gain ~ pups_born_alive, data = .) %>%
  broom::tidy()
```

`wt_gain` outcome 
`pups_born_alive` 
KEY: period sign `.` 





